\documentclass[12pt]{article}

% Packages
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[backend=biber,style=authoryear,citestyle=authoryear,natbib=true]{biblatex}
\addbibresource{references.bib}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{titlesec}
\setlength{\parskip}{1em} 
\setlength{\bibitemsep}{1em} % Add spacing between bibliography entries

%Customizing section titles
% Customize section and subsection titles
\titleformat{\section}
  {\normalfont\Large\bfseries} % Section titles will be in large, bold font
  {\thesection}{1em}{}
  [\titlerule]

\titleformat{\subsection}
  {\normalfont\bfseries} % Subsection titles will be slightly smaller
  {\thesubsection}{1em}{}
\titlespacing{\subsection}{0pt}{*0}{-0.1em}

% Title
\title{Adaptation Plan Evaluation Report}
\author{Mukund Balaji Srinivas | u7274095}
\date{} % Remove the date 

\begin{document}

% Title Page
\maketitle

\begin{abstract}
This section provides a brief summary of the evaluation report, outlining the key question being addressed, methods used, and main conclusions.
\end{abstract}

\newpage

\section*{Introduction}
% \subsection*{Background}

The World Economic Forum (\cite{masterson_2024})  outlines six key technologies critical for climate change adaptation, 
including the powerful roles of Earth Observation [EO] and Artificial Intelligence [AI]. 


\subsection*{Earth Observation}
EO sector plays a pivotal role in climate change adaptation by offering critical data for monitoring deforestation, land-use changes, and rising sea levels (\cite{anderson_2017}). 
The EO technologies are essential for tracking progress towards the 2030 Agenda for Sustainable Development (\cite{unitednations_2015}), particularly in providing spatially explicit data to monitor targets across 
a wide range of SDGs [Sustainable Development Goals]. EO supports informed decision-making by governments and international 
organizations, allowing for better resource management and monitoring of the impact of climate mitigation strategies.
However, as EO becomes more industrialized (\cite{venkatesan_2020}) , it presents several significant environmental risks  
that must be considered when evaluating its contribution to climate adaptation (\cite{crisp_2020}). \\
While EO transitions to a more cloud-based data processing model, it introduces significant environmental 
considerations within the environmental science community. Storing and distributing EO data relies heavily on a global network of data centers, 
which are known to contribute substantially to carbon emissions and disproportionately use water resources. Despite these impacts, the specific 
environmental footprint of EO data centers remains opaque and requires critical assessment alongside its contributions to climate adaptation.(\cite{wilkinson_2024})

\subsection*{Artificial Intelligence}
AI models require immense computational power and their energy demands continue to grow exponentially (\cite{amodei_2018}),
 their potential contribution to carbon emissions becomes a significant concern (\cite{taddeo_2021}). Nonetheless, leveraging AI’s 
 expanding capabilities to improve climate change mitigation is both feasible and crucial. 
 For instance, AI can be used to better understand the impacts of industrial production, construction, 
 and transportation on climate change, while also estimating the effectiveness of policy interventions such 
 as carbon taxes and carbon trading systems. However, ensuring the sustainable and equitable deployment of these 
 AI-driven solutions will require robust governance frameworks (\cite{taddeo_2021}).
%AI models require massive computational power and their energy demands grow exponentially (\cite{amodei_2018}), having the potential to significantly contribute to carbon emissions (\cite{Henderson_2020}).




At the same time , the  Information and Communication Technology [ICT] sector that encompasses EO and AI
is responsible for over 3 percent of the world's carbon emissions (\cite{jones_2018}) . Specifically , 
companies like Samsung and Amazon had a footprint of around 40,000MT of \(CO_2\) equivalent in 2021 (\cite{navarro_2023_the}) —raises the question 
of how big tech companies are addressing their environmental impact. While AI and EO tools offer potential benefits in 
mitigating climate change, their carbon costs are often overlooked. 

Given the dual role of ICT as both a significant contributor to global carbon emissions and a 
key driver of climate change solutions, it is crucial to examine the adaptation strategies 
adopted by tech companies to mitigate their environmental impact. This report systematically 
evaluates the sustainability reports of major tech companies—Meta, Apple, Amazon, Google, Nvidia, 
Samsung and Dell—focusing on their key businesses, associated climate risks, and summaries of 
their adaptation strategies. Currently, no specific 
framework exists to evaluate these strategies in the tech sector, necessitating reliance on established climate change 
metrics. To this effect, the techniques outlined in the framework provided by the International Institute for Environment and Development 
will guide the evaluation, ensuring that these companies’ adaptation policies align with broader sustainability goals. The evaluation will utilize 
the scorecard suggested by the IIED (\cite{craft_2016}) for systematic reporting, along with the framework proposed in the article by Jones N. (\cite{jones_2001}) 
to assess the adaptation strategies of the selected companies effectively.

\pagebreak
\section*{Methods}
Earth Observation [EO] plays a vital role in achieving the Sustainable Development Goals (SDGs) outlined in the 2030 Agenda. EO technologies help track, monitor, 
and report progress toward these goals by enabling data-driven decision-making, improving the accuracy of national statistics through spatially explicit data, and 
addressing cross-cutting themes such as climate and energy. The Group on Earth Observations (GEO) is pivotal in ensuring EO's practical application in 
supporting the 2030 Agenda, particularly through partnerships, as emphasized in Strategic Development Goal 17. EO's contributions to national planning and fostering 
synergy between various development sectors can significantly enhance the achievement of SDG targets, providing a strong return on investment.


\section*{Comparison}
Here, make the necessary comparisons based on the criteria from the Methods section. Discuss how different nations/regions/sectors perform according to these criteria. You may break this into subsections if there are multiple comparisons.

\section*{Discussion}
Synthesize the information you gathered from the comparisons. Discuss patterns, insights, and important findings based on the comparisons made.

\section*{Conclusion}
Summarize the main conclusions of your evaluation. Reflect on what the comparisons suggest about the criteria and any recommendations based on the results.

\newpage
\printbibliography

\end{document}

%Changes more 'K